---
title: "Heart Disease Indicators"
author: "Kyle Hollands"
date: "26/04/2022"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r scripts, echo = FALSE, include=FALSE, message=FALSE, warning=FALSE}
# Load the required libraries and set options
library(tidyverse)
library(janitor)
library(readr)
library(tidyr)
library(chron)
library(dplyr)
library(readr)
library(lubridate)
library(ggridges)
options("scipen"=10)

# Import the dataset
heart_health_metrics <- read.csv("./Datasets/Modified CSV/heart.csv")

# Rename some of the columns.
heart_health_metrics <- rename(heart_health_metrics,
  RestingBloodPressure = RestingBP,
  FastingBloodSugar = FastingBS,
  MaxHeartRate = MaxHR)

# Remove any rows where Cholesterol and Resting Blood Pressure equates to 0.
heart_health_metrics <- heart_health_metrics[!(heart_health_metrics$Cholesterol <= 0),]
heart_health_metrics <- heart_health_metrics[!(heart_health_metrics$RestingBloodPressure <= 0),]

# Create some additional labels for the charts.
(heart_disease_labels <- c(
                    `0` = "Normal",
                    `1` = "Heart Disease",
                    `F` = "Female",
                    `M` = "Male"
                    ))
```

## Table of Contents

- Objective
- The Data Analyzed
- Visualizations
- Conclusion
- Recommendations

## Objective

- To indicate what factors may contribute to heart disease.

## The Data Analyzed
- In all analysis, the common contributors of Heart Disease are included, such as Cholesterol, Age and Blood Pressure.
    - The data indicates that there does appear to be a connection between higher Cholesterol, Age and Blood Pressure with a greater risk of Heart Disease.

- Other metrics analyzed:
    - Pre-diabetics and diabetics
    - ST-T wave abnormality
    - ECG wave types
    - Exercise angina
    - Chest pain type
    - A lower max heart rate

## Data
- Heart Disease and Cholesterol
```{r data_1, echo=FALSE ,warning=FALSE,message=FALSE}
setNames(aggregate(heart_health_metrics$Cholesterol ~ heart_health_metrics$HeartDisease, FUN = mean), c("Cholesterol","Heart Disease"))
```
- Heart Disease and Diabetes
```{r data_2, echo=FALSE ,warning=FALSE,message=FALSE}
setNames(aggregate(heart_health_metrics$FastingBloodSugar ~ heart_health_metrics$HeartDisease, FUN = mean), c("Diabetes","Heart Disease"))
```
- Heart Disease and Max Heart Rate
```{r data_2, echo=FALSE ,warning=FALSE,message=FALSE}
setNames(aggregate(heart_health_metrics$MaxHeartRate ~ heart_health_metrics$HeartDisease, FUN = mean), c("Max Heart Rate","Heart Disease"))
```
## Visualizations

## Diabetes and Heart Disease
```{r chart_1, echo=FALSE, results='hide',fig.keep='all',warning=FALSE,message=FALSE}
ggplot(heart_health_metrics, aes(x = Age, y = Cholesterol)) +
  geom_density_2d(alpha = 0.50, size = 0.1) +
  theme(legend.position = "bottom") +
  geom_point(aes(color = FastingBloodSugar)) +
  geom_smooth(method="loess", se=F, aes(x = Age, y = RestingBloodPressure)) +
  facet_grid(. ~ HeartDisease, labeller = as_labeller(heart_disease_labels)) +
  ylim(75,600) +
  scale_color_viridis_c() +
  #guides(color = "none") +
  labs(
    x = "Age", y = "Cholesterol (mm/dl)",
    title = "Does Diabetes contribute to Heart Disease?",
    subtitle = "1 = Pre-diabetes/diabetes | Blue line = Blood Pressure",
    caption = "Source: Kaggle - Heart Failure Prediction Dataset"
  )
```

## ST-T Wave Abnormality and Heart Disease
```{r chart_2, echo=FALSE, results='hide',fig.keep='all',warning=FALSE,message=FALSE}
ggplot(heart_health_metrics, aes(x = Age, y = Cholesterol)) +
  geom_density_2d(alpha = 0.50, size = 0.1) +
  theme(legend.position = "bottom") +
  geom_point(aes(color = Oldpeak)) +
  geom_smooth(method="loess", se=F, aes(x = Age, y = RestingBloodPressure)) +
  facet_grid(. ~ HeartDisease, labeller = as_labeller(heart_disease_labels)) +
  ylim(75,600) +
  scale_color_viridis_c() +
  #guides(color = "none") +
  labs(
    x = "Age", y = "Cholesterol (mm/dl)",
    title = "Does an ST-T wave abnormality contribute to Heart Disease?",
    subtitle = "Lower value = less risk | Blue line = Blood Pressure",
    caption = "Source: Kaggle - Heart Failure Prediction Dataset"
  )
```

## ECG Wave Types and Heart Disease
```{r chart_3, echo=FALSE, results='hide',fig.keep='all',warning=FALSE,message=FALSE}
ggplot(heart_health_metrics, aes(x = Age, y = Cholesterol)) +
  geom_density_2d(alpha = 0.50, size = 0.1) +
  theme(legend.position = "bottom") +
  geom_point(aes(color = ST_Slope)) +
  geom_smooth(method="loess", se=F, aes(x = Age, y = RestingBloodPressure)) +
  facet_grid(. ~ HeartDisease, labeller = as_labeller(heart_disease_labels)) +
  ylim(75,600) +
  scale_color_viridis_d() +
  labs(
    x = "Age", y = "Cholesterol (mm/dl)",
    title = "What ECG wave types contribute to a higher risk of Heart Disease?",
    subtitle = "Blue line = Blood Pressure",
    caption = "Source: Kaggle - Heart Failure Prediction Dataset"
  )
```

## Exercise Angina and Heart Disease
```{r chart_4, echo=FALSE, results='hide',fig.keep='all',warning=FALSE,message=FALSE}
ggplot(heart_health_metrics, aes(x = Age, y = Cholesterol)) +
  geom_density_2d(alpha = 0.50, size = 0.1) +
  theme(legend.position = "bottom") +
  geom_point(aes(color = ExerciseAngina)) +
  geom_smooth(method="loess", se=F, aes(x = Age, y = RestingBloodPressure)) +
  facet_grid(. ~ HeartDisease, labeller = as_labeller(heart_disease_labels)) +
  ylim(75,600) +
  scale_color_viridis_c() +
  #guides(color = "none") +
  labs(
    x = "Age", y = "Cholesterol (mm/dl)",
    title = "Does Exercise Angina contribute to Heart Disease?",
    subtitle = "1 = Exercise Angina | Blue line = Blood Pressure",
    caption = "Source: Kaggle - Heart Failure Prediction Dataset"
  )
```

## Chest Pain Type vs Heart Disease

```{r chart_5, echo=FALSE, results='hide',fig.keep='all',warning=FALSE,message=FALSE}
ggplot(heart_health_metrics, aes(x = Age, y = Cholesterol)) +
  geom_density_2d(alpha = 0.50, size = 0.1) +
  theme(legend.position = "bottom") +
  geom_point(aes(color = ChestPainType)) +
  geom_smooth(method="loess", se=F, aes(x = Age, y = RestingBloodPressure)) +
  facet_grid(. ~ HeartDisease, labeller = as_labeller(heart_disease_labels)) +
  ylim(75,600) +
  scale_color_viridis_d() +
  labs(
    x = "Age", y = "Cholesterol (mm/dl)",
    title = "What Chest Pain types contribute to a higher risk of Heart Disease?",
    subtitle = "Blue line = Blood Pressure",
    caption = "Source: Kaggle - Heart Failure Prediction Dataset"
  )
```


## Max Heart Rate and Heart Disease
```{r chart_6, echo=FALSE, results='hide',fig.keep='all',warning=FALSE,message=FALSE}
ggplot(heart_health_metrics, aes(x = Age, y = Cholesterol)) +
  geom_density_2d(alpha = 0.50, size = 0.1) +
  theme(legend.position = "bottom") +
  geom_point(aes(color = MaxHeartRate)) +
  geom_smooth(method="loess", se=F, aes(x = Age, y = RestingBloodPressure)) +
  facet_grid(. ~ HeartDisease, labeller = as_labeller(heart_disease_labels)) +
  ylim(75,600) +
  scale_color_viridis_c() +
  #guides(color = "none") +
  labs(
    x = "Age", y = "Cholesterol (mm/dl)",
    title = "Does a higher max heart rate contribute to Heart Disease?",
    subtitle = "Blue line = Blood Pressure",
    caption = "Source: Kaggle - Heart Failure Prediction Dataset"
  )
```

## Conclusion
- In addition to Cholesterol, Age and Blood Pressure, the following appear to lead to a greater risk of Heart Diease:
    - Pre-diabetes / diabetes
    - A higher ST-T wave abnormality value
    - Flat and down ECG wave types
    - Exercise angina
    - Asymptomatic chest pain
    - A lower max heart rate

## Recommendations
- Certain lifestyle changes, such as diet and exercise could possibly help mitigate some, if not all, risk factors of Heart Disease to a 
certain extent. However, this is not professional medical advice and one should consult their healthcare practitoner.