---
title: "Heart Disease Indicators"
author: "Kyle Hollands"
date: "26/04/2022"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Table of Contents

- Objective
- What the data shows
    - Diabetes and Heart Disease
    - ECG type and Heart Disease
    - Exercise Angina and Heart Disease
    - Chest Pain Type and Heart Disease
    - Max Heart Rate and Heart Disease
- Conclusion
- Recommendations

## Objective

- To indicate what factors contribute to Heart Disease, and with this information provide data-driven recommendations on how to mitigate it.

## What the data shows
- In all analysis, the common contributors of Heart Disease are included, such as Cholesterol, Age and Blood Pressure.
    - The data indicates that there does appear to be a connection between higher Cholesterol, Age and Blood Pressure with a greater risk of Heart Disease.

- The remaining metrics that appear to lead to a greater risk:
    - Pre-diabetic / diabetic.
    - Horizontal or downsloping ECG wave type.
    - Exercise angina.
    - Asymptomatic chest pain type.
    - A lower max heart rate.

## Visualizations

## Diabetes and Heart Disease

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

